### 2024-10-27 main_widom_morse.tex

\section{Introduction}
A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point. Below the critical point the substance exhibits a liquid-gas coexistence. At the critical point, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state. Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like structures under supercritical conditions, and display a unique mix of properties. With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power. Moreover, the supercritical properties are instantly tunable by changing pressure and temperature. This is the reason why supercritical fluids have diverse applications across various industries \cite{KnezEtAl2014}.

In the context of supercritical fluids, the Widom line, Frenkel line, and percolation line are theoretical and physical constructs used to describe the complex behavior of fluids beyond the critical point.
All three concepts provide a way to describe how fluids behave in the supercritical region, where the traditional liquid-gas phase distinction no longer applies.
The Frenkel line \cite{BrykEtAl2017} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion). The percolation line \cite{CampiEtAl2001} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.

In the present work we are particularly interested in the Widom line \cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions. Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XuEtAl2005}, is treated as the continuation of the coexistence line above the critical point.   
Some thermodynamic response functions, for example, isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region. In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{ArtemenkoEtAl2017} and then diverge from each other as the temperature rise. Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor2020Book}.

@@

@@

In the present work we take advantage of the results obtained in \cite{KozlovskiiDobush2020,DobushEtAl2024}. In \cite{KozlovskiiDobush2020}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system. In the context of liquid-vapour systems, these collective variables are linked to modes of density deviation from the critical density. This set of collective variables encompasses those which average values are connected to order parameters.Thus, the depiction of the phase transition within the phase space of collective variables appears compelling.} 
is used for calculating the grand partition function of a cell fluid model in the zero-mode approximation and consequently deriving the explicit forms of the equation of state (EoS) both in pressure-temperature-chemical potential terms and pressure-temperature-density terms. The resulting EoSs are applicable in a wide range of temperatures below and above the critical point. In \cite{DobushEtAl2024}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.

@@

The potential of interaction between particles is taken in a form of a modified Morse potential
\begin{eqnarray}
	\label{def:mod_morse}
	U(r) & = & \varepsilon C_{H} \left[A {\rm e}^{-n_0(r-R_0)/\alpha} \right.
	\nonumber\\
	&& + \left. {\rm e}^{-\gamma(r - R_0)/\alpha} -2{\rm e}^{-(r-R_0)/\alpha}\right],
\end{eqnarray}
where $R_0$ is the coordinate of the potential minimum, $\alpha$ is an effective range of interaction, $\gamma$ and $n_0$ are parameters of the model. Other two constants $C_{H}$ and $A$ are expressed via $\gamma$ and $n_0$ as follows
\begin{equation}
	C_{H} = \frac{n_0}{n_0 + \gamma - 2}, \quad A = \frac{2 - \gamma}{n_0},
\end{equation}
where $\varepsilon$ is the depth of the potential well at $r=R_0$. Applying $\gamma=2$ reduce the  to the ordinary Morse potential~\cite{Morse1929} at $\gamma=2$. For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KozlovskiiDobush2020}.

Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KozlovskiiDobush2016}.

@@

The equation of state obtained in~\cite{KozlovskiiDobush2020} reads
\begin{equation}
	\label{eq:eosMT}
	Pv\beta = E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4.
\end{equation}
The quantities in the left-hand side of the equation are $P$, the pressure; $\beta = (k_{\rm B} T)^{-1}$, the inverse temperature; $k_{\rm B}$, the Boltzmann constant; $T$, the temperature; $v$, cell volume. The quantities in the right-hand side are, in general, functions of the temperature $T$ and the chemical potential $\mu$. Let us present their expressions explicitly.

@@

The parameter $\chi_0$ is used in~\cite{KozlovskiiDobush2020} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KozlovskiiDobush2020}.

@@

The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KozlovskiiDobush2020} it was found that
\begin{equation*}
	\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
\end{equation*}
We also use the reduced temperature defined as $T^* = k_{\rm B}T/\varepsilon$, and thus its critical value $T^*_c = 4.995.$

Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KozlovskiiDobush2020}:
\begin{eqnarray}
	\label{params}
	\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
	n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
	\alpha^* = 5.0 & \quad p = 1.0.
\end{eqnarray}

The quantity $d$ entering equations~\eqref{eq:eosMT} and~\eqref{chem_pot} is a function of temperature
\begin{equation}
	\label{def:D0}
	d = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)}.
\end{equation}
The condition $d = 0$ defines the critical temperature~\cite{KozlovskiiDobush2020}
\begin{equation}
	k_{\rm B}T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
\end{equation}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$ at $T \geq T_c$ represented by black lines. Grey lines correspond to isotherms of pressure at $T < T_c$ based on the results for the same model taken from~\cite{KozlovskiiDobush2020}. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), іsotherms are curves consisting of two segments connected by a vertex (grey lines in Figure~\ref{fig1}). This form indicates a first-order phase transition occurs in the system. Within our approach~\cite{KozlovskiiDobush2020} at $T<T_c$, negative values of the effective chemical potential mean the system is in a gaseous state. The vertex $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much sharper. At the critical temperature pressure is a smooth plane curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, since below the critical temperature, $M=0$ stands for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse} based on the equation of state~\eqref{eq:eosPTM_reduced}. 
	}\label{fig2}
\end{figure}

@@

 In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas. Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary. Nevertheless, fluids exhibit distinct liquid-like and gas-like structures beyond the critical temperature. A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010}. This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances. Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,  in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}). It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
 In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DobushEtAl2024} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020,DobushEtAl2024}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, derive the isothermal compressibility in explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}
\begin{figure}[h!] 
	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The two figures differ in the scale of $M$. Part~(a) covers a wider range of $M$. Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}
Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$). Right at the critical point the compressibility diverge. As a result, there is still a "memory" of the critical point, where $\kappa_T$ reach local maxima, but once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases. 
Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities. Above the critical point, the fluid becomes continuous. However, Figure~\ref{fig3}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region. In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior. In the second case ($M>0$), it behaves like a liquid. Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$. At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig3}b), they completely vanish at $\tau > 0.28$.

To identify the Widom line maximize the isothermal compressibility by solving the equation
\begin{equation}\label{eq:maximize}
	\frac{\partial \kappa_T^*}{\partial M} = 0.
\end{equation}
Applying the solution of \eqref{eq:maximize} in the equation of state \eqref{def:reduced_pres}, one locates the Widom line position in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.

@@

@@

>>>>

\section{Introduction}
A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point. Below the critical point the substance exhibits a liquid-gas coexistence. At the critical point, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state. Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like behavior under supercritical conditions, and display a unique mix of properties. With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power. Moreover, the supercritical properties are instantly tunable by changing pressure and temperature. This is the reason why supercritical fluids have diverse applications across various industries \cite{KMLetal14}.

In the context of supercritical fluids, the Widom line, Frenkel line, and percolation line are theoretical and physical constructs used to describe the complex behavior of fluids beyond the critical point.
All three concepts provide a way to describe how fluids behave in the supercritical region, where the traditional liquid-gas phase distinction no longer applies.
The Frenkel line \cite{BGMetal17} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion). The percolation line \cite{CKS01} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.

In the present work we are particularly interested in the Widom line \cite{XKBetal05,MS10,AKM17} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions. Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XKBetal05}, is treated as the continuation of the coexistence line above the critical point.   
Some thermodynamic response functions, for example, isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region. In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{AKM17} and then diverge from each other as the temperature rise. Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor20}.

@@

[Here a short paragraph that this work is a logical consequence of our previous works on the equation of state and response functions for the cell fluid model with interaction expressed by the modified Morse potential]

@@

In the present work we take advantage of the results obtained in \cite{KD20,DKRP24arxiv,DKRP24}. In \cite{KD20}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system. In the context of liquid-vapour systems, these collective variables are linked to modes of density deviation from the critical density. This set of collective variables encompasses those which average values are connected to order parameters.Thus, the depiction of the phase transition within the phase space of collective variables appears compelling.} 
is used for calculating the grand partition function of a cell fluid model in the zero-mode approximation and consequently deriving the explicit forms of the equation of state (EoS) both in pressure-temperature-chemical potential terms and pressure-temperature-density terms. The resulting EoSs are applicable in a wide range of temperatures below and above the critical point. In \cite{DKRP24arxiv,DKRP24}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.

@@

The potential of interaction between particles is taken in a form of a modified Morse potential
\begin{eqnarray}
	\label{def:mod_morse}
	U(r) & = & \varepsilon C_{H} \left[A {\rm e}^{-n_0(r-R_0)/\alpha} \right.
	\nonumber\\
	&& + \left. {\rm e}^{-\gamma(r - R_0)/\alpha} -2{\rm e}^{-(r-R_0)/\alpha}\right],
\end{eqnarray}
where $R_0$ is the coordinate of the potential minimum, $\alpha$ is an effective range of interaction, $\gamma$ and $n_0$ are parameters of the model. Other two constants $C_{H}$ and $A$ are expressed via $\gamma$ and $n_0$ as follows
\begin{equation}
	C_{H} = \frac{n_0}{n_0 + \gamma - 2}, \quad A = \frac{2 - \gamma}{n_0},
\end{equation}
where $\varepsilon$ is the depth of the potential well at $r=R_0$. Applying $\gamma=2$ reduces $U(r)$ to the ordinary Morse potential~\cite{Morse29}. For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}.

Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KD16}.

@@

The equation of state obtained in~\cite{KD20} reads
\begin{equation}
	\label{eq:eosMT}
	Pv\beta = E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4.
\end{equation}
The quantities in the left-hand side of the equation are $P$, the pressure; $\beta = (k_{\rm B} T)^{-1}$, the inverse temperature; $k_{\rm B}$, the Boltzmann constant; $T$, the temperature; $v$, cell volume. The quantities in the right-hand side are, in general, functions of the temperature $T$ and the chemical potential $\mu$. Let us present their expressions explicitly.

@@

The parameter $\chi_0$ is used in~\cite{KD20} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KD20}.

@@

The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KD20} it was found that
\begin{equation*}
	\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
\end{equation*}
We also use the reduced temperature defined as $T^* = k_{\rm B}T/\varepsilon$, and thus its critical value $T^*_c = 4.995.$

Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KD20}:
\begin{eqnarray}
	\label{params}
	\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
	n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
	\alpha^* = 5.0 & \quad p = 1.0.
\end{eqnarray}

The quantity $d$ entering equations~\eqref{eq:eosMT} and~\eqref{chem_pot} is a function of temperature
\begin{equation}
	\label{def:D0}
	d = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)}.
\end{equation}
The condition $d = 0$ defines the critical temperature~\cite{KD20}
\begin{equation}
	k_{\rm B}T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
\end{equation}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (grey lines in Figure~\ref{fig1}). This form indicates a first-order phase transition in the system. Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ mean the system is in a gaseous state. The condition $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much sharper. At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, since below the critical temperature, $M=0$ is the condition for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosPTM_reduced}. 
	}\label{fig2}
\end{figure}

@@

 In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas. Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary. Nevertheless, fluids exhibit distinct liquid-like and gas-like behavior beyond the critical temperature. A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}. This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances. Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,  in~\cite{DKRP24arxiv,DKRP24} we established that the isothermal compressibility of the cell fluid model diverge in the critical point, and consequently it displays maxima in the supercritical region (see Figures~\ref{fig3}). It means the possibility to derive the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
 In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KD20,DKRP24arxiv,DKRP24}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, derive the isothermal compressibility in explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}
\begin{figure}[h!] 
	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The two figures differ in the scale of $M$. Part~(a) covers a wider range of $M$. Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}
Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$). Right at the critical point the compressibility diverges. As a result, there is still a "memory" of the critical point, where $\kappa_T$ reaches local maxima. However, once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases. 
Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities. Above the critical point, the fluid becomes continuous. However, Figure~\ref{fig3}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region. In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior. In the second case ($M>0$), it behaves like a liquid. Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$. At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig3}b), they completely vanish at $\tau > 0.28$.

To identify the Widom line, the locus of the isothermal compressibility maxima is found by solving the equation
\begin{equation}\label{eq:maximize}
	\frac{\partial \kappa_T^*}{\partial M} = 0
\end{equation}
with respect to $M$. 
Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{def:reduced_pres}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.

@@

%\bibliographystyle{elsarticle-num}
\bibliographystyle{plain}
\bibliography{extracted}

@@

%\end{document}
